---
title: "Using Keeper with Large Language Models"
author: "Martijn Schuemie"
date: "`r Sys.Date()`"
output:
  pdf_document:
    number_sections: yes
    toc: yes
  html_document:
    number_sections: yes
    toc: yes
vignette: >
  %\VignetteIndexEntry{Using Keeper with LLMs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(Keeper)
```

# Introduction

This vignette describes how one could use Keeper to generate patient summaries, and have them be reviewed by a large-language model (LLM).
This vignette assumes you have read and understand the 'Generating Keeper profiles for review' vignette. 
We will build on the same example used in that other vignette: Type 1 Diabetes Mellitus (T1DM), and assume you have the various objects created in that vignette.

# LLM review

We assume the Keeper input concept sets have already been defined, and that Keeper has been executed:

```{r echo=FALSE,message=FALSE,eval=TRUE}
keeper <- readRDS(system.file("shuffledKeeper.rds", package = "Keeper"))
```
```{r}
keeper
```

We use the `ellmer` package to connect to an LLM from your provider of choice, including Anthropic, Google, OpenAI, or a local LLM. For example, we can connect to OpenAI's ChatGPT using:

```{r eval = FALSE}
library(ellmer)
client <- chat_openai()
```

This assumes you have set the `OPENAI_API_KEY` environmental variable. See the [ellmer package](https://ellmer.tidyverse.org) for details on connecting to various providers.
Next, we can have the LLM review the cases:

```{r eval = FALSE}
promptSettings <- createPromptSettings()
llmResponses <- reviewCases(keeper = keeper,
                            settings = promptSettings,
                            phenotypeName = "Type I Diabetes Mellitus (T1DM)",
                            client = client,
                            cacheFolder = "cache")
```
```{r echo=FALSE,eval=TRUE}
message("Reviewing person 1 of 20\n...\nReviewing person 20 of 20\nReviewing cases took 3.9 mins and cost $NA")
```

Here, we use the default prompt settings. 
The phenotype name is what is inserted in the LLM prompt, so must match the exact disease you wish to adjudicate. 
The `cacheFolder` is a location on your file system where the prompts and responses will be saved. 
If for some reason the process stops early, you can rerun and responses that have already been generated will be skipped.

```{r echo=FALSE,message=FALSE,eval=TRUE}
llmResponses <- readRDS(system.file("llmResponses.rds", package = "Keeper"))
```
```{r}
llmResponses
```

From here, it is trivial to compute the positive predictive value (PPV) of our cohort:

```{r}
mean(llmResponses$isCase == "yes")
```

# Using a highly-sensitive cohort to estimate PPV and sensitivity

Instead of starting with an existing cohort for which we wish to compute the PPV, we can also create a highly-sensitive cohort that is unlikely to miss any true cases. 
When we use LLMs to review a large sample of such a cohort we will be able to compute not only PPV but also sensitivity. 

## Creating a highly sensitive cohort


The `createSensitiveCohort()` automatically function constructs the highly sensitive cohort based on the input concept sets that are also used by the `generateKeeper()` function.
See the `createSensitiveCohort()` function documentation for more details.

```{r eval = FALSE}
createSensitiveCohort(
  connectionDetails = connectionDetails,
  cdmDatabaseSchema = cdmDatabaseSchema,
  cohortDatabaseSchema = cohortDatabaseSchema,
  cohortTable = cohortTable,
  cohortDefinitionId = 2,
  createCohortTable = FALSE,
  keeperConceptSets = conceptSets
)
```

Here, `conceptSets` is the collection of concept sets described in the 'Generating Keeper profiles for review' vignette. 
We use the same cohort table as that other vignette, so we assume we do not need to create it.
We use `cohortDefinitionId = 2` so as not to conflict with the cohort definition with ID 1 we created earlier.

## Run Keeper on the highly-sensitive cohort

Next we run Keeper on the highly-sensitive cohort. 

```{r eval = FALSE}
keeperHsc <- generateKeeper(
  connectionDetails = connectionDetails,
  cohortDatabaseSchema = cohortDatabaseSchema,
  cdmDatabaseSchema = cdmDatabaseSchema,
  cohortTable = cohortTable,
  cohortDefinitionId = 2,
  sampleSize = 10000,
  phenotypeName = "T1DM",
  keeperConceptSets = conceptSets,
  removePii = FALSE
)
```

We take a large sample (10,000 person) to ensure we have enough positive cases to compute PPV and sensitivity.
Note that we specify `removePii = FALSE` so we can later join the LLM adjudications with specific cohorts we wish to evaluate.

## LLM Review of the highly-sensitive cohort

Next we have the LLM review all 10,000 sampled patient profiles.
Be aware that this can incur significant costs. 
For example, having our preferred model (GPT o3) review 10,000 cases typically costs between $60-$80.

```{r eval = FALSE}
promptSettings <- createPromptSettings()
llmResponsesHsc <- reviewCases(keeper = keeperHsc,
                               settings = promptSettings,
                               phenotypeName = "Type I Diabetes Mellitus (T1DM)",
                               client = client,
                               cacheFolder = "cacheVignetteHsc")
```

## Uploading the reference cohort

## Evaluating cohorts using the reference cohort



